Algebraic File Synchronization: Adequacy and Completeness

### 2018-07-10 p2/conclusion.tex
Minor improvements
Added conclusion

Apart from constructing a filesystem-free algebra,
there are many other ways in which further work can extend the current results.
An important extension would be to
consider reconciling not only two, but more replicas in a single step and
prove the correctness and maximality of the algorithm proposed,
or show that it is impossible to satisfy these criteria.

@@

Future work could also investigate extending the 
model and algorithms to the $|\setd|>1$ case
so that directory metadata could be represented directly
as opposed to through pre- and post-processing as described
in \cref{sec_multidir}.

>>>>

In this paper we presented an algebraic model to file synchronization,
and defined an update detector and a reconciliation algorithm.
With its extended command set the model was carefully designed to take
advantage of an inherent symmetry of filesystem commands,
which was enhanced by removing meta-information from directories
as well as the \emph{move} command.
These assumptions enabled us to prove that
the update detector algorithm is correct (\cref{update_works}),
and that reconciliation is not only correct (\cref{reconciliation_correct})
but also yields maximal results (\cref{rec_is_complete}).

We introduced relations and concepts that have proven particularly
useful, including 
independent commands, % (\cref{def_indep}), 
simple sequences, % (\cref{def_min_simp}),
and relationships between domains of sets of command sequences ($\worksmeqsign$)
that allowed us to formalize the correctness of reconciliation.
Among our partial results were interesting properties of these concepts,
for example,
that all legal permutations of simple sequences are equivalent (\cref{simple_reorder_equiv}),
that command sequences can be transformed into simple sequences using syntactical rules (\cref{can_simplify}),
or that the combination of independent sequences is defined wherever the components are defined (\cref{combine_independent_sequences}).
We believe these results will prove useful outside their current usage as well.

Our results can be immediately applied to practical filesystem synchronizers,
which could benefit from e.g. recording changes using the extended command set,
and from the theoretical background that guarantees their correct operation.
Furthermore, with the encoding and decoding steps described in \cref{sec_multidir},
the reconciliation algorithm can be extended to handle directory meta-information
and the \emph{move} command,
which makes our model applicable to more use cases and filesystems.

\myskip
Apart from constructing a filesystem-free algebra,
there are many other ways in which further work can extend the current results.
An important extension would be to
consider reconciling not only two, but more replicas in a single step and
prove the correctness and maximality of the algorithm proposed,
or show that it is impossible to satisfy these criteria.

@@

Future work could also investigate extending the 
model and algorithms to the $|\setd|>1$ case
so that directory metadata could be represented directly
as opposed to through encoding and decoding steps.

### 2018-07-10 p2/def-commands.tex
Added conclusion

\myskip
For reasons also listed in \cite{NREC}, in this model we will not consider
a \textit{move} or \textit{rename} command.
This turns out to be useful because this would be the only command that affects
filesystems at two nodes at once, therefore describing 
the dependencies for \textit{move} would call for a more complicated model.
As \cref{sec_multidir} details, this restriction does not mean that in an application
implementing conflict resolution using the algorithm described here would not be
able to handle renames by pre- and post-processing changes in the filesystem to
discover them, which can enhance usability.

>>>>

\myskip
For reasons also listed in \cite{NREC}, in this model we will not consider
a \textit{move} or \textit{rename} command.
This turns out to be useful because this would be the only command that affects
filesystems at two nodes at once, therefore describing 
the dependencies for \textit{move} would call for a more complicated model.
As \cref{sec_multidir} details, this restriction does not mean that in an application
implementing conflict resolution using the algorithm described here would not be
able to handle renames by post-processing changes in the filesystem to
discover them, which can enhance usability.

### 2018-07-10 p2/intro.tex
Spellcheck
Added conclusion

The approach presented in this paper offers an improvement
over results desribed in \cite{NREC} in multiple ways.
We introduce a new set of commands that is symmetric and
captures more information about the updates,
and we restrict information content in directories
to exploit a further hidden symmetry in the command set.
These have the effect that reasoning about commands
becomes simpler as there are fewer edge cases,
and more powerful as 
predictions made by the reasoning are more accurate
due to the additional information content.
In fact, the new command set not only simplifies
our reconciliation algorithm, but also makes it maximal.
\cite{NREC} also lacked proofs that the update detection
and reconciliation algorithms it presented work as intended,
which we provide in the current paper.
While the results are intuitive, providing rigorous proofs
is far from trivial.
During the process, we define a number of useful concepts
and show their relationships and the properties they possess.
In our view these construct a special algebraic model
that is worthy of interest and of further research on its own.

The paper is organised as follows.
%
We start by defining a model of filesystems and a set of commands
that we will use to describe updates and modifications in \cref{sec_def}.
%
This is followed by investigating the properties and behaviors
of command pairs in \cref{section_axioms}.
%
\Cref{sec_update} describes update detection and ordering filesystem commands.
The main result of this section if that under some simple conditions,
a set of filesystem commands executed in any feasible order leads
to the same changes in the filesystem.
%
The reconciliation and conflict detection algorithm is defined in
\cref{sec_rec}, which merges two sets of commands.
We then proceed to prove that the output of reconciliation
is correct in the sense that 
it can be applied to the replicas without causing an error,
and it is maximal
inasmuch as no further updates can be applied under any circumstances.
In order to be able to do this, we introduce
domains of sets of command sequences,
and show that it has a number of highly convenient properties.
%
\Cref{sec_multidir,sec_algebra,sec_conclusion}
describe extending our model to widen its applicability;
outline directions for further work on the introduced algebraic system,\
and provide the conclusion.
%
We discuss related work, comparing our results to other research in \cref{sec_relatedwork},
while
\cref{axiom_proof} contains technical proofs for propositions on the behavior of command pairs.

>>>>

The approach presented in this paper offers an improvement
over results described in \cite{NREC} in multiple ways.
We introduce a new set of commands that is symmetric and
captures more information about the updates,
and we restrict information content in directories
to exploit a further hidden symmetry in the command set.
These have the effect that reasoning about commands
becomes simpler as there are fewer edge cases,
and more powerful as 
predictions made by the reasoning are more accurate
due to the additional information content.
In fact, the new command set not only simplifies
our reconciliation algorithm, but also makes it maximal.
\cite{NREC} also lacked proofs that the update detection
and reconciliation algorithms it presented work as intended,
which we provide in the current paper.
While the results are intuitive, providing rigorous proofs
is far from trivial.
During the process, we define a number of useful concepts
and show their relationships and the properties they possess.
In our view these construct a special algebraic model
that is worthy of interest and of further research on its own.

The paper is organized as follows.
%
We start by defining a model of filesystems and a set of commands
that we will use to describe updates and modifications in \cref{sec_def}.
%
This is followed by investigating the properties and behaviors
of command pairs in \cref{section_axioms}.
%
\Cref{sec_update} describes update detection and ordering filesystem commands.
The main result of this section if that under some simple conditions,
a set of filesystem commands executed in any feasible order leads
to the same changes in the filesystem.
%
The reconciliation and conflict detection algorithm is defined in
\cref{sec_rec}, which merges two sets of commands.
We then proceed to prove that the output of reconciliation
is correct in the sense that 
it can be applied to the replicas without causing an error,
and it is maximal
inasmuch as no further updates can be applied under any circumstances.
In order to be able to do this, we introduce
domains of sets of command sequences,
and show that it has a number of highly convenient properties.
%
\Cref{sec_multidir,sec_algebra,sec_conclusion}
describe extending our model to widen its applicability;
outline directions for further work on the introduced algebraic system,
and provide the conclusion.
%
We discuss related work, comparing our results to other research in \cref{sec_relatedwork},
while
\cref{axiom_proof} contains technical proofs for propositions on the behavior of command pairs.

### 2018-07-10 p2/multidir.tex
Minor improvements
Added conclusion

The two assumptions we used to increase the symmetry of our model were
that there is no move command,
and that
there is only one directory value, that is, directories are not differentiated
by the meta-information they contain.
In this section we describe reasons why the latter assumption may not limit the applicability
of even the current model, and we also describe pre- and post-processing steps
to overcome the two restrictions.

We note that commercial designs for synchronizers 
often avoid considering metadata in directories as
as these are generally not understood well by users,
and, if needed, conflict resolution on these settings can be easily automated.
See, for example, the seminal work of Balasubramaniam and Pierce and the Unison synchronizer \cite{BP} and \cite{BZ}.

>>>>

The two assumptions we used to increase the symmetry of our model were
that there is no move command,
and that
there is only one directory value, that is, directories are not differentiated
by the meta-information they contain.
In this section we describe reasons why the latter assumption may not limit the applicability
of even the current model, and we also describe encoding and decoding steps
to overcome the two restrictions.

We note that commercial designs for synchronizers 
often avoid considering metadata in directories as
as these are generally not understood well by users,
and, if needed, conflict resolution on these settings can be easily automated.
See, for example, \cite{BZ} and the seminal work of Balasubramaniam and Pierce and the Unison synchronizer \cite{BP}.

### 2018-07-10 p2/upd-def.tex
Added conclusion

\begin{mydef}[Minimal and simple command sets and command sequences]
A sequence or set of commands is \emph{minimal} if it contains at most one command on each node.
It is \emph{simple} if it is minimal and it does not contain assertion commands.
\end{mydef}

>>>>

\begin{mydef}[Minimal and simple command sets and command sequences]\mlabel{def_min_simp}
A sequence or set of commands is \emph{minimal} if it contains at most one command on each node.
It is \emph{simple} if it is minimal and it does not contain assertion commands.
\end{mydef}
